---
title: 'STAT 231: R Questions & Answers'
author: "Evan Daisy, Kim Zhou, Sanjana Sunder"
date: "Feb. 25, 2021"
output: pdf_document
---

```{r, setup, include=FALSE}
# Load packages here 
library(mdsr)   
library(babynames)
library(tidyverse)

knitr::opts_chunk$set(
  tidy=FALSE,     # display code as typed
  size="small",
  message=FALSE)   # slightly smaller font for code
```

# Original code

Below is the code, slightly modified, from the extended example on historical baby names in section 3.3.1.  You can work off the section(s) of this code relevant to your assigned question as you explore answers.

```{r, include=FALSE}
# "The 'lifetables' table from the 'babynames' package contains actuarial estimates of the number of people per 100,000 who are alive at age x, for every 0 <= x <= 114.  The 'make_babynames_dist()' function in the 'mdsr' package adds some more convenient variables and filters for only the data that is relevant to people alive in 2014" -- MDSR pg. 49
head(lifetables, 5)

BabynamesDist <- make_babynames_dist()
head(BabynamesDist, 5)

joseph <- BabynamesDist %>%
  filter(name == "Joseph" & sex == "M")

name_plot <- ggplot(data = joseph, aes(x = year)) +
  geom_bar(stat = "identity", aes(y = count_thousands*alive_prob)
           , fill = "#b2d7e9", color = "white") + 
  geom_line(aes(y = count_thousands), size = 2) + 
  ylab("Number of People (thousands)") +
  xlab(NULL) 
  
name_plot

wtd.quantile <- Hmisc::wtd.quantile
median_yob <- with(joseph, wtd.quantile(year, est_alive_today, probs = 0.5))
median_yob

# Figure 3.22
name_plot2 <- name_plot + 
  geom_bar(stat = "identity", color = "white", fill = "#008fd5"
           , aes(y = ifelse(year == median_yob, est_alive_today/1000, 0))) +
  ggtitle("Age Distribution of American Boys Named Joseph") +
  geom_text(x = 1935, y = 40, label = "Number of Josephs \nborn each year") +
  geom_text(x = 1915, y = 13
            , label = "Number of Josephs \nborn each year \nestimated to be alive \non 1/1/2014"
            , color = "#b2d7e9") +
  geom_text(x = 2003, y = 40, label = "The median \nliving Joseph \nis 37 years old"
            , color = "darkgray") + 
  theme_dark() +
  geom_curve(x = 1995, xend = 1974, y = 40, yend = 24
             , arrow = arrow(length = unit(0.3,"cm")), curvature = 0.5) +
  ylim(0,42)

name_plot2

# Figure 3.23
name_plot %+% filter(BabynamesDist, name == "Josephine" & sex == "F")

# Figure 3.24
names_plot <- name_plot + facet_wrap(~sex) 
names_plot %+% filter(BabynamesDist, name == "Jessie")

# Figure 3.25
many_names_plot <- name_plot + facet_grid(name ~ sex)
many_names_plot %+% filter(BabynamesDist, name %in% c("Jessie", "Marion", "Jackie"))

# Figure 3.27 (Most common women's names)
com_fem <- BabynamesDist %>%
  filter(sex == "F" & est_alive_today > 0) %>%
  group_by(name) %>%
  summarise(
    N = n()
    , est_num_alive = sum(est_alive_today)
    , q1_age = wtd.quantile(age_today, est_alive_today, probs = 0.25)
    , median_age = wtd.quantile(age_today, est_alive_today, probs = 0.5)
    , q3_age = wtd.quantile(age_today, est_alive_today, probs = 0.75)) %>%
  arrange(desc(est_num_alive)) %>%
  head(25)

w_plot <- ggplot(data = com_fem
                 , aes(x = reorder(name,-median_age), y = median_age)) + 
  xlab(NULL) +
  ylab("Age (in years)") +
  ggtitle("Median ages for females with the 25 most common names") +
  geom_linerange(aes(ymin = q1_age, ymax = q3_age), color = "#f3d478"
                 , size = 10, alpha = 0.8) +
  geom_point(fill = "#ed3324", color = "white", size = 4, shape = 21) +
  coord_flip()

w_plot
```



### Q6. What else can be done with the geom_text function (options for rotation, size, font, color, etc.)?  I see there are coordinates to choose where to display the text, but was wondering how you are supposed to know the grid dimensions? 

Include your answer and explanation here.  You can intersperse text with code in your explanation.
###The geom_text function is used to generate text on the plot or graphic. It helps with the design and format of the overall display. Within the geom_text function we can change the position of the text using x and y = to specify where the text is centered. Additionally, we can change the size of the label/text by typing size = #. Using the geom_text function we can also adjust fonts with the aes() nested in and rotate the text within the aes() function as well. Finally, we can change the color of the text generated in the geom_text function using color= and then a common color or its hex#. In terms of grid-lines, these are generated separate from the geom_text function, using the theme_dark(), for example, for a darker grey background.
```{r}
# geom_text(x = 1935, y = 40, label = "Number of Josephs \nborn each year", aes(family=c("serif")), size = 2, angle = 90)


```

Once you're satisfied with your explanation:

- set the r code chunk option "include=FALSE" for the code chunk above (for the original code); we don't want to see all the original code in your answer and that option will suppress that code and output from being printed (but will still evaluate the code)
- delete the text and section header for the "# Original Code" section
- knit the document to pdf
- go to the Google group conversation “R Questions & Answers from PS1A”, and “Reply all”;  include the question you are answering in the body of the message, and attach your knit PDF with the response

 